%% LaTeX_Thesis_Template.tex
% An unofficial LaTeX template for Cranfield theses.

%%
% This document is an example of the use of the unofficial "cranfieldthesis" 
% LaTeX style file.  I hope it's useful, and good like.

%% Options for the bibliography generator
%% \bib_opts[bib_file]{library.bib}
%% \bib_opts[output_file]{bibliography.tex}

\documentclass[12pt, oneside]{book}

\usepackage{luatex85}

\usepackage[pdftex, pdfpagelabels=true, linktocpage, hidelinks]{hyperref}
% \usepackage{hypcap}

% Use the custom "cranfieldthesis" LaTeX style file. 
\usepackage{cranfieldthesis}
\usepackage{amssymb}
\usepackage{geometry}
\usepackage[titletoc]{appendix}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{lscape}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{titletoc}
\usepackage{tikz}
\usepackage[bottom]{footmisc}
\usepackage{float}
\usetikzlibrary{shapes, arrows, matrix}
\usepackage{adjustbox}
\usepackage{cellspace}
\setlength\cellspacetoplimit{4pt}
\setlength\cellspacebottomlimit{4pt}


\usepackage{scrlayer}

\input{utilities}

% Example parameters for a typical taught MSc course
\title{A Web Based Terrain Modeller using Fractals and CAD}
\author{Miguel Marques}
\date{August 2016}
\school{Aerospace, Transport and Manufacturing}
\course{Computational and Software Techniques in Engineering \\ Digital Signal and Image Processing}
\degree{MSc}
\academicyear{2015--2016}
\supervisor{Dr Peter Sherar}
\copyrightyear{2016}


\begin{document}


%% Front matter
%
% This is where we do the title page, etc.
%

\frontmatter

% Standard-Form Title Pages
\maketitle

\pagestyle{plain}

% Abstract and Keywords
\begin{abstract}
	\doublespacing
	
    The process of modelling man made objects is, classically, a manual process where a user deforms a 3D object though the use of several techniques. This process is time-consuming and, with the usage of procedural techniques, can be improved, enabling the creation of more realistic and detailed content. One such case where this techniques have been widely used is in the creation of virtual terrains, not only because of the realistic details, but also due to the possibility of creating huge random infinite terrains that are used in the video game industry.
    Procedural generation has the disadvantage of loss of detailed control over the modelled terrain, that is, using a purely procedural method it is complicated to specify where the major features of the terrain should be positioned. This is the main problem that this thesis addresses.
    
    In this project a hybrid terrain modelling process is proposed. In this process the user creates a simplified version of the desired terrain containing the basic topology, which is then blended with a random generated surface to create a detailed version of the terrain using image processing techniques. Additionally a web application, where this process is implemented, was developed using \textit{Angular.js} and \mbox{\textit{WebGL}}. The architecture for this application is detailed in this thesis. In order to improve the performance of the random generation and the blending process some GPU computations were implemented using \mbox{\textit{WebGL 2.0}}. This computations were then benchmarked and the comparison with their CPU counterparts is shown in this document.
    
    

	\subsubsection*{Keywords}
	Procedural Generation; Terrain Modelling; Web Application; WebGL; Fractal geometry; GPU; Fourier filtering; Noise Synthesis
\end{abstract}

% Acknowledgements
\chapter{Acknowledgements}

\begin{notes}
  \item The author would like to thank
\end{notes}

% Table of Contents
\sstableofcontents

% List of Figures
\sslistoffigures

% List of Tables
\sslistoftables

% List of Equations
\sslistofequations

% The list of abbreviations can't be automatically generated so you need to populate it yourself
\begin{listofabbreviations}
    \abbrev{fBm}{Fractional Brownian Motion}
%    \abbrev{PMHCI}{Picewise Monotone Hermite Cubic Interpolation}
    \abbrev{HPF}{High-pass Filter}
    \abbrev{FFT}{Fast Fourier Transform}
    \abbrev{IFFT}{Inverse Fast Fourier Transform}
\end{listofabbreviations}


%% Main Matter
%
% This is where we include the main thesis content.
%
\mainmatter

\include{introduction}
\include{literature_review}
\include{methodology}
\include{software_architecture}
\include{results}
\include{conclusion}

% References - you can use BiBTeX inside the 'references' environment if you want.
% For short reference lists, you might find it just as easy to use the bibitem form.
\begin{references}
	\input{bibliography}
\end{references}

%% Back matter
%
% This is where we include appendices and references

\chapter*{Appendices}
\addcontentsline{toc}{chapter}{Appendices}

\begin{appendices}	
	\input {user_manual}
	\input {additional_results}
\end{appendices}

\backmatter

\end{document}

